---
title: "Assignment 4"
author: "Raquel Luna"
editor: visual
format:
  html:
    page-layout: article
    progress: true
    scrollable: true
---

## Charts 1-4 for Cost of Living data set

```{r, echo=FALSE, warning=FALSE}
# Load necessary library
library(ggplot2)

# Load the dataset (assuming the file is in the same directory as your Quarto file)
cost.of.living <- read.csv("cost-of-living.csv")

# Use the correct column names for Houston and Dallas
cities <- c("Houston..TX..United.States", "Dallas..TX..United.States")
meal_cost <- cost.of.living[3, c("Houston..TX..United.States", "Dallas..TX..United.States")]  # Extract "Meal for 2" costs
market_share <- c(1.5, 1.5)  # Adjust market share to prevent overlap

# Create a data frame for comparison
data <- data.frame(City = cities, MealCost = as.numeric(meal_cost), MarketShare = market_share)

# Create a variable-width column chart with different colors for each city
ggplot(data, aes(x = City, y = MealCost, width = MarketShare, fill = City)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Dallas..TX..United.States" = "steelblue", "Houston..TX..United.States" = "coral")) +  # Custom colors
  theme_minimal() +
  labs(title = "Meal Cost for 2 in Houston and Dallas", x = "City", y = "Cost (USD)")

```

Meal Cost for Two in Houston and Dallas This chart illustrates the cost of a mid-range, three-course meal for two people in Houston, TX and Dallas, TX. The data is based on current cost-of-living figures, highlighting the price differences between these two major cities in Texas.

Houston has a slightly higher average cost of \$53.91 for a meal for two, while Dallas shows a lower cost at \$44.92. The width of each bar represents an arbitrary "market share" in this chart to demonstrate variable-width column charts, though this could represent the size of the dining market or the popularity of restaurants in these cities. The contrasting colors make it easy to distinguish between the two cities and quickly compare the meal prices. ---

```{r, echo=FALSE, warning=FALSE}
# Load necessary library
library(ggplot2)

# Load the dataset (assuming the file is in the same directory as your Quarto file)
cost.of.living <- read.csv("cost-of-living.csv")

# Select multiple cities to compare (for example)
cities <- c("Houston..TX..United.States", "Dallas..TX..United.States", "Austin..TX..United.States", "Chicago..IL..United.States")
meal_cost <- cost.of.living[3, c("Houston..TX..United.States", "Dallas..TX..United.States", "Austin..TX..United.States", "Chicago..IL..United.States")]  # Extract "Meal for 2" costs

# Create a data frame for comparison
data <- data.frame(City = cities, MealCost = as.numeric(meal_cost))

# Create a bar chart to compare meal costs across cities with rotated x-axis labels
ggplot(data, aes(x = City, y = MealCost, fill = City)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Houston..TX..United.States" = "coral", 
                               "Dallas..TX..United.States" = "steelblue", 
                               "Austin..TX..United.States" = "green", 
                               "Chicago..IL..United.States" = "purple")) +  # Custom colors for each city
  theme_minimal() +
  labs(title = "Meal Cost for 2 in Selected US Cities", x = "City", y = "Cost (USD)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels

```

Meal Cost Comparison Across Selected US Cities This bar chart compares the cost of a mid-range, three-course meal for two people across four selected cities in the United States: Houston, Dallas, Austin, and Chicago. The data highlights differences in meal pricing across these urban areas:

Houston has the highest meal cost at approximately \$53.91. Dallas follows closely with a meal cost of \$44.92. Austin and Chicago both have meal costs around \$50, slightly below that of Houston. The custom colors and rotated x-axis labels provide an easy-to-read comparison of meal costs in these cities, helping to visualize regional differences in cost of living. ---

```{r, echo=FALSE, warning=FALSE}
# Load necessary library
library(ggplot2)

# Load the dataset (assuming the file is in the same directory as your Quarto file)
cost.of.living <- read.csv("cost-of-living.csv")

# Select multiple categories to compare for a single city (Houston, TX)
categories <- c("Meal for 2", "1-bedroom Apartment", "Gasoline (per liter)")
costs <- cost.of.living[c(3, 24, 18), "Houston..TX..United.States"]  # Select rows for Meal cost, Apartment Rent, and Gasoline price

# Create a data frame for comparison
data <- data.frame(Category = categories, Cost = as.numeric(costs))

# Create a bar chart comparing costs for different categories in Houston with a logarithmic y-scale
ggplot(data, aes(x = Category, y = Cost, fill = Category)) +
  geom_bar(stat = "identity") +
  scale_fill_manual(values = c("Meal for 2" = "lightblue", 
                               "1-bedroom Apartment" = "coral", 
                               "Gasoline (per liter)" = "green")) +  # Custom colors for each category
  scale_y_log10() +  # Use logarithmic scale to handle large differences
  theme_minimal() +
  labs(title = "Cost Comparison for Different Categories in Houston", x = "Category", y = "Cost (USD, Log Scale)") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels for readability
```

Cost Comparison for Different Categories in Houston This bar chart compares the costs of three key categories in Houston, Texas: Meal for 2, 1-bedroom Apartment Rent, and Gasoline (per liter). Due to the large disparity in values between these categories, the chart uses a logarithmic scale on the y-axis to provide a clearer comparison:

The cost of renting a 1-bedroom apartment in Houston is significantly higher, around \$2,311.84. The cost of a mid-range meal for two people is approximately \$53.91, much lower in comparison. The price of gasoline per liter is around \$1.17. ---

```{r, echo=FALSE, warning=FALSE}
# Load necessary library
library(ggplot2)

# Load the dataset (assuming the file is in the same directory as your Quarto file)
cost.of.living <- read.csv("cost-of-living.csv")

# Select categories and costs for Houston and Dallas
categories <- c("Meal for 2", "Gasoline (per liter)", "1-bedroom Apartment")
houston_cost <- cost.of.living[c(3, 18, 24), "Houston..TX..United.States"]
dallas_cost <- cost.of.living[c(3, 18, 24), "Dallas..TX..United.States"]

# Create a data frame for comparison
data <- data.frame(
  City = rep(c("Houston", "Dallas"), each = length(categories)),
  Category = rep(categories, times = 2),
  Cost = c(houston_cost, dallas_cost)
)

# Create a faceted column chart
ggplot(data, aes(x = City, y = Cost, fill = City)) +
  geom_col() +
  facet_wrap(~ Category, scales = "free") +  # Facet by category and free scales
  scale_fill_manual(values = c("Houston" = "coral", "Dallas" = "steelblue")) +  # Custom colors for each city
  theme_minimal() +
  labs(title = "Cost Comparison for Different Categories in Houston and Dallas", x = "City", y = "Cost (USD)")

```

Cost Comparison for Different Categories in Houston and Dallas This faceted chart compares the costs of three different categories—Meal for 2, Gasoline (per liter), and 1-bedroom Apartment—in Houston and Dallas. The chart is split into separate panels for each category, making it easier to compare the cost differences between the two cities.

The 1-bedroom Apartment rent is the most expensive category in both cities, with Dallas being slightly higher than Houston. The cost of Gasoline per liter and Meal for 2 is relatively similar in both cities, but still shows minor differences. By breaking the data into different facets, the chart allows for better visibility and comparison between categories that vary greatly in cost. ---
